---
title: "Lesson 3"
author: "Haozhi"
date: "2022-11-03"
output: html_document
---
# Loop functions
Row is observation, column is variabls.

### something new
* `range(..., na.rm = FALSE)` returns the minimum and maximum of its first argument, which should be a numeric vector.\
* `unique(x, incomparables = FALSE, ...)` returns a vector, data frame or array like x but with duplicate elements/rows removed.\
* To pass a newly defined function (anonymous functions) in another function, `{}` is not required if this newly-defined function takes only one argument. like `lapply(x, function(elem) elem[2])`, elem is just a dummy variable.\
* `table(flags$landmass)` will return a summary table which sum each factors of landmass in the flags matrix.\
* `rnorm(n, mean = 0, sd = 1)` normal distribution.\
* `runif(n, mean = 0, sd = 1)` uniform distribution, or `runif(n, min = 0, max = 1)` to add the \
* `str()` can be used to generate the structure of a function.\
* `rep` is the replicate.
* `gl(n, k, length = n * k, labels = seq_len(n), ordered = FALSE)` to generate factor by n levels and k replicates in each level.\
* `with(data, expr, ...)` to call out the parent environment for expr evaluation. like 
`tapply(mtcars$mpg, mtcars$cyl, mean)` can be simplified as `with(mtcars, tapply(mpg, cyl, mean))`




### `apply`
```{r}
str(apply) #(X, MARGIN = (1 or 2), FUN, ...)
```
* evaluate the function over the margins (dimensions like row = 1 or column = 2) of an array.\
* But `Colsum`, `Colmean`, `Rowsum`, `Rowmean` will normally do the same with much faster speed.\
* will return a vector by default because `simplify = TRUE` argument.\

### `lapply`
```{r}
str(lapply) 
```

* easier than for or while loop when programming in a compact command line.\
* take s list as input, applies a function to each element (column) of the list, and always return a list of the same length as the original\
* l stand for list, and no parentheses for the function inside all loop functions.\

### `sapply`
```{r}
str(sapply)
```

* simplified version of `lapply`\
* if the result is a list where every element is of length one, then sapply() returns a vector. If the result is a list where every element is a vector of the same length (> 1), sapply()returns a matrix. If sapply() can't figure things out, then it just returns a list, no different from what lapply() would give you.\

### `mplay`
```{r}
str(mplay)
```
* multivariate apply of sorts which applies a function in parallel over a set of arguments.\
* easy to use when there is vector input, example:`mapply(rep, 1:4, 4:1)`, `mapply(mean, 1:5, 1:5)`

### `vapply`
```{r}
str(vapply)
```
* vapply() allows to specify the result format explicitly, If the result doesn't match the format you specify, vapply() will throw an error, causing the operation to stop.\
* safer than sapply function, and perform faster than sapply for large datasets.\

### `tapply`
```{r}
str(tapply) 
```
* split the data into groups based on the value of some variable (factors), then apply function to the members of each group.\
* `tapply(flags$animate, flags$landmass, mean)` think about how this works!\

### `split`
```{r}
str(split)
```
* takes a vector or other objects (x) and splits it into groups determined by a factor or list of factors (f), drop to remove empty values.\

### Some examples
* tapply(mtcars$mpg, mtcars$cyl, mean), with(mtcars, tapply(mpg, cyl, mean)), 
sapply(split(mtcars$mpg, mtcars$cyl), mean) are the same.\


### Debugging
* `message`, `warning`, `error`, and only error is fatal.\
* check list:\
  + what was your input? how did you call the function?\
  + what were you expecting? output, message, other results?\
  + what did you get?\
  + how does what you get differ from what you were expecting?\
  + Were your expectations correct in the first place?\
  + Can you reproduce the problem(exactly)?\
* Tools:\
  + `traceback()` to print out the function call stack after an error occurs.\
  + `debug()`put a function for "debug" mode to step through the execution one line at a time.\
  + `browser()`: suspends the execution.\
  + `trace()` only works with the latest code to trace back where error occurred, and `recover()`.\
  
















